% sections/03_glossary.tex

\section*{Glossary}
\label{sec:glossary}

This glossary lists the entities used in the document. Within the model
universe, nothing exists beyond these items, their Distinctions, and the
Relations they support~\cite{reginald2025af}.

\begin{description}

  \item[Null State ($\mathcal{N}$)]
    Relational starting point. No Distinctions and no Relations are present.
    Symbols $\infty$ and $\chi$ are used synonymously with $\mathcal{N}$ in
    this document.

  \item[Distinction]
    Primitive difference used to describe structure. Any structured object in
    this document is described as a finite family of Distinctions.

  \item[Relation]
    Connection between Distinctions. Relations have no independent existence;
    they are entirely determined by, and exist only through, Distinctions.

  \item[Region]
    Grouped family of positions, represented by a subset $R \subseteq V$ of an
    index set $V$ for Distinctions. A Region induces an inside/outside split,
    and a boundary given by those ordered pairs $(i,j)$ with $i \in R$ and
    $j \notin R$.

  \item[Memory State]
    Probability assignment on a finite family of Distinctions. A Memory State
    is a function $p : V \to [0,1]$ with $\sum_{i\in V} p(i) = 1$.

  \item[Observer / Modeller]
    Name for the Memory State and its updates when a reference label is
    required. No separate object beyond the Memory State and the Choices for
    its update is assumed.

  \item[Markov Kernel]
    Matrix $P = (P_{ij})_{i,j\in V}$ of transition probabilities on a finite
    family of Distinctions, with $P_{ij} \ge 0$ and $\sum_{j} P_{ij} = 1$ for
    each $i \in V$. Together, $(p,P)$ specify how the Memory State changes
    between updates.

  \item[Flux ($\phi_{ij}$)]
    Quantity defined from the Memory State $p$ and the Markov Kernel $P$ on
    ordered pairs of indices $i,j\in V$. In this document, net flux is
    antisymmetric:
    \[
      \phi_{ij} := p(i) P_{ij} - p(j) P_{ji},
      \qquad
      \phi_{ij} = -\phi_{ji}.
    \]

  \item[Divergence ($\mathrm{div}$)]
    Sum of flux at an index,
    \[
      \mathrm{div}(i) := \sum_{j\in V} \phi_{ij}.
    \]

  \item[AFC Stokes Identity]
    Identity relating an interior sum of divergence to boundary flux. For any
    Region $R \subseteq V$,
    \[
      \sum_{i\in R} \mathrm{div}(i)
      =
      \sum_{\substack{i\in R \\ j\notin R}} \phi_{ij},
    \]
    with $\phi$ and $\mathrm{div}$ as above and boundary defined from $R$.

  \item[Capacity]
    Information-theoretic quantity associated with a Markov Kernel $P$, given
    by the channel capacity
    \[
      C(P) := \max_{p} I_p(X;Y),
    \]
    where $I_p(X;Y)$ is the mutual information between successive Memory
    States induced by $(p,P)$ and the maximum is taken over all Memory States
    on $V$.
\end{description}
